Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-05-11 ltx/data_paper/DataPaper.tex
edits to Result section

\subsection{Estimating Support For Connectivity Kernel}
Under the assumption that the sensors are not spatially band-limiting the spectral content of the field and the connectivity kernel is homogeneous, the support for the connectivity kernel can be inferred by studying the spatial cross-correlation between consecutive observations. The deterministic component of the spatial mapping between consecutive fields is due to the convolution of the connectivity kernel with the firing rates. Although the relationship between the field and the firing rate is nonlinear, the cross-correlation analysis still extracts meaningful results. This can be seen by considering the three operating regions of the sigmoid. When the mean membrane voltage is below the active region of the sigmoid, the firing rate will be approximately zero and only very small spatial correlations will observed. Therefore, this will not contribute to the results. Within the active region the sigmoid can be approximated by a linear function between consecutive samples, given a sufficiently high sampling rate. Therefore, linear cross-correlation analysis is appropriate. When the sigmoid is in the saturated region, the correlation coefficients will be higher, but not provide a meaningful measure of the kernel weight or gain. Therefore, we can not estimate the gain for the kernel using cross-correlation analysis, but the spatial extent can still be inferred. The gain for the kernel will be inferred using the EM algorithm described below.

@@

\section{Results}
This section demonstrates the identification of the proposed IDE model using simulated data. Data is generated employing state space equation (\ref{eq:StateEq}-\ref{eq:QmatrixForSigmapoints}) using a square grid of $n_x= 9 \times 9$ equally spaced ($2.5~mm$) basis functions with $\sigma_{\phi}^2=1.58~mm^2$  , assuming the observed field can be fully
described by the decomposed IDE model. 100 realizations of 500 ms of data were generated and the first 100 ms was discarded allowing the model's dynamic to reach steady state. The observation noise was set to $\Sigma_{\epsilon}=0.2\times \mathbf I_{n_y}$ and the field disturbance was set to $\mathcal{GP}(\mathbf 0,0.1 \times\gamma(\mathbf{r}-\mathbf{r}'))$ where the covariance function was chosen as a Gaussian with zero mean and the width $\sigma_{\gamma}^2=1.7~mm^2$ . The distance between adjacent sensors was 1.5 mm resulting into $n_y = 196$ observations. The width of the sensor kernels were set to $\sigma_m^2=0.81~mm^2$. The sampling period and the synaptic time constant were set to $T_s = 1~ms$  and $\tau = 10~ms$, respectively. The sigmoidal activation parameters were $\varsigma=0.56~mV^{-1}$ and $v_0=1.8~mV$.
\begin{itemize}
	\item Will show parameter distributions
	\item Will show lower bound
\end{itemize}

@@

\begin{itemize}
	\item comment on how the e step is UKF and M-step is kinda like the EKF, and we dont use EKF in E-step since UKF gives better estimates and the linearisation in M-step would be more accurate. 
	\item comment on EM has been used with EKS by Ghahramani which has poorer convergance property (I guess) using EM with URTSS
\end{itemize}

% % all figures here
% % ~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelWidthEstimation.pdf}
\end{center}
\caption{{\bf Connectivity kernel support estimation}.}
\label{fig:KernelWidthEstimation}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelBasis.pdf}
\end{center}
\caption{{\bf Connectivity kernel basis functions, $n_{\theta}=25$}, $\sigma^2_{\theta}=3.5$.}
\label{fig:KernelBasis}
\end{figure}
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelBasis1.pdf}
\end{center}
\caption{{\bf Connectivity kernel basis functions, $n_{\theta}=25$}, $\sigma^2_{\theta}=3.5$.}
\label{fig:KernelBasis1}
\end{figure}
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/DisturbanceWidthEstimation.pdf}
\end{center}
\caption{{\bf Disturbance support estimation}.}
\label{fig:DisturbanceWidthEstimation}
\end{figure}

>>>>

\subsection{Correlation analysis}\label{subsec:CorrAnal}
Under the assumption that the sensors are not spatially band-limiting the spectral content of the field and the connectivity kernel is homogeneous, the support for the connectivity kernel can be inferred by studying the spatial cross-correlation between consecutive observations. The deterministic component of the spatial mapping between consecutive fields is due to the convolution of the connectivity kernel with the firing rates. Although the relationship between the field and the firing rate is nonlinear, the cross-correlation analysis still extracts meaningful results. This can be seen by considering the three operating regions of the sigmoid. When the mean membrane voltage is below the active region of the sigmoid, the firing rate will be approximately zero and only very small spatial correlations will observed. Therefore, this will not contribute to the results. Within the active region the sigmoid can be approximated by a linear function between consecutive samples, given a sufficiently high sampling rate. Therefore, linear cross-correlation analysis is appropriate. When the sigmoid is in the saturated region, the correlation coefficients will be higher, but not provide a meaningful measure of the kernel weight or gain. Therefore, we can not estimate the gain for the kernel using cross-correlation analysis, but the spatial extent can still be inferred. The gain for the kernel will be inferred using the EM algorithm described below.

@@

\section{Results}
This section demonstrates the identification of the proposed IDE model using simulated data. The observation noise and disturbance were unknown and no assumption were made on the shape of the kernel.  Data was generated employing state space equation (\ref{eq:StateEq}-\ref{eq:QmatrixForSigmapoints}) using a square grid of $n_x= 9 \times 9$ equally spaced ($2.5~mm$) basis functions with $\sigma_{\phi}^2=1.58~mm^2$, assuming the observed field can be fully described by the decomposed IDE model. 100 realizations of 500 ms of data were generated and the first 100 ms was discarded allowing the model's dynamic to reach steady state. The observation noise was set to $\Sigma_{\epsilon}=0.2\times \mathbf I_{n_y}$ and the field disturbance was set to $\mathcal{GP}(\mathbf 0,0.1 \times\gamma(\mathbf{r}-\mathbf{r}'))$ where the covariance function was chosen as a Gaussian with zero mean and the width $\sigma_{\gamma}^2=1.7~mm^2$ . The distance between adjacent sensors was 1.5 mm resulting into $n_y = 196$ observations. The width of the sensor kernels were set to $\sigma_m^2=0.81~mm^2$. The sampling period and the synaptic time constant were set to $T_s = 1~ms$  and $\tau = 10~ms$, respectively. The sigmoidal activation parameters were $\varsigma=0.56~mV^{-1}$ and $v_0=1.8~mV$.

The correlation analysis described in Section \ref{subsec:CorrAnal} were applied to generated data in order to approximate the unknown supports of both connectivity kernel and the disturbance covariance function. In the algorithm, the unknown variables $\sigma_{\epsilon}^2$ and $\xi$ in equation () and () were set to zero in correlation analysis. By doing so, the exact shape of the kernel and disturbance covariance function cannot be inferred, however, an approximate of their support can be still extracted. In fact, the estimated support for the disturbance covariance function was very accurate. The estimates of the connectivity kernel and disturbance support were normalised and thresholded for values less than 0.01. The result of the correlation analysis for the connectivity kernel is shown in Figure.\ref{fig:KernelWidthEstimation}. The red circle shows the estimated support which encompasses the non-zero values after thresholding. The estimated support of the disturbance covariance function convolved twice with the observation kernel is shown in Figure.\ref{fig:DisturbanceWidthEstimation}. By fitting a Guassian basis function to the result, and under the assumption that the width of the sensor,$\sigma_m^2$ were known, the width of the disturbance covariance function, $\sigma_d^2$ was calculated analytically using the expression for Gaussian basis functions convolution.

\begin{itemize}
	\item Will show parameter distributions
	\item Will show lower bound
      \end{itemize}

@@

\begin{itemize}
	\item comment on how the e step is UKF and M-step is kinda like the EKF, and we dont use EKF in E-step since UKF gives better estimates and the linearisation in M-step would be more accurate. 
	\item comment on EM has been used with EKS by Ghahramani which has poorer convergance property (I guess) using EM with URTSS
	\item itterative EM and correlation analysis (maybe)
\end{itemize}

% % all figures here
% % ~~~~~~~~~~~~~~~~

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelBasis1.pdf}
\end{center}
\caption{{\bf Connectivity kernel basis functions, $n_{\theta}=25$}, $\sigma^2_{\theta}=3.5$.}
\label{fig:KernelWidthEstimation}
\end{figure}
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/DisturbanceWidthEstimation.pdf}
\end{center}
\caption{{\bf Disturbance support estimation}.}
\label{fig:DisturbanceWidthEstimation}
\end{figure}
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelWidthEstimation.pdf}
\end{center}
\caption{{\bf Connectivity kernel support estimation}.}
\label{fig:KernelWidthEstimation1}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelBasis.pdf}
\end{center}
\caption{{\bf Connectivity kernel basis functions, $n_{\theta}=25$}, $\sigma^2_{\theta}=3.5$.}
\label{fig:KernelBasis}
\end{figure}
